sensiAnPaper1All

### 2016-02-09 paper1_results_figs.Rmd
new figures

---
  title: "effectOfIndividualInputs"
  output: pdf_document
  author: "Andy South"
  date: "`r Sys.Date()`"
---

# Investigation into effects of individual inputs on time to reach resistance thresholds.
# version 2 - in progress

## Based on 1000 runs with all inputs varying at once which explains the variability at particular input values.
## Within each figure the plots are divided into 6 sub-plots according to the insecticide use strategy and when the resistance threshold is reached
1. insecticide 1 : sole use
1. insecticide 2 : sole use
1. Mixture 1 : threshold reached for either insecticide in mixture
1. Mixture 2 : once threshold reached for either insecticide in mixture, switch to sole use of other until it too reaches threshold
1. Mixture 3 : threshold reached for both insecticides in mixture
1. Sequential : sole use of one insecticide, switch to other when threshold reached until it too reaches threshold

These plots show gen_cP0.2, which is the number of generations to reach 20% resistance.

Red dashed lines are a smoothed mean.

Runs where resistance thresholds have not been reached have been removed.

Previous plots showed a set of points at 1000 generations. The model is run for 500 generations, any runs which have not reached the resistance threshold by this time are given a value of 1000. This has little effect on questions of whether mixtures or sequential use is better.

@@

  ## trying with the extended experiment _ex100
  experiment <- 'extended'
  load(file=paste0(outFolder,'listOutMix_ex1000.rda'))
  load(file=paste0(outFolder,'listOutI1_ex1000.rda'))
  load(file=paste0(outFolder,'listOutI2_ex1000.rda'))

@@

```{r,echo=FALSE,results='hide'}
### chunk copied from sensiAnPaper1All
### now want to calculate the times to reach critical points and add them on to the input file 
### for different insecticide strategies, sequential, mix1 and mix2

  #1) sequential : time to resistance for each insecticide in isolation
  #inputs : inAndOutI1, inAndOutI2
  #find time to criticalPoint for insecticide1
  #find time to criticalPoint for insecticide2
  #add together
  resistPointsI1 <- findResistancePoints(listOutI1, locus=1)
  resistPointsI2 <- findResistancePoints(listOutI2, locus=2)

  resistPointsSeq <- resistPointsI1 + resistPointsI2

@@

  #subsetting now done above before names prettified
  #treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','effectiveness_insecticide1','effectiveness_insecticide2','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
  #treeInput <- treeInput[ treePredictors, ]

@@

## This plot shows exposure which has the most noticeable pattern.
```{r, eval=TRUE, echo=FALSE, warning=FALSE}

@@

  #function to transpose and add strategy column with the passed value
  #also add on the input columns  
  addStrategyColumn <- function(x, value, inputs){
    x <- as.data.frame( t(x) )
    x$strategy <- value
    #transpose inputs
    inputs <- as.data.frame( t(inputs) )
    #cbind onto the outputs
    x <- cbind(inputs,x)
    x
  }

@@

  #ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + facet_wrap( ~ cut)

  #cool working!

  #exposure has the most noticeable effect
  ggplot(ggInsOuts, aes(x=exposure, y=time_to_resistance0.25, color=strategy)) + 
    geom_point(shape=3) + 
    facet_wrap( ~ strategy) + 
    #geom_smooth(aes(x=exposure, y=gen_cP0.5, color=NULL))
    geom_smooth(colour='red', linetype='dashed',size=0.5)

  #effectiveness of insecticide1 little noticeable effect
#   ggplot(ggInsOuts, aes(x=effectiveness_insecticide1, y=time_to_resistance0.25, color=strategy)) + 
#     geom_point() + 
#     facet_wrap( ~ strategy) + 
#     #geom_smooth(aes(x=effectiveness_insecticide1, y=gen_cP0.5, color=NULL)) 
#     geom_smooth(colour='red', linetype='dashed',size=0.5)

@@

## In the following plots all inputs are plotted individually, the points are coloured by exposure, which allows us to see that, even as other inputs change, low times to resistance are associated with high exposure.
```{r, eval=TRUE, echo=FALSE, fig.height=3.5}

@@

@@

    y <- 'time_to_resistance0.25'    
    #y <- 'time_to_resistance0.5'
    #color <- 'strategy'
    #coloring by exposure shows its effect even when other params are varying
    color <- 'exposure'

    print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
             geom_point(shape=3) + 
             facet_wrap( ~ strategy) + 
             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
             geom_smooth(colour='red', linetype='dashed',size=0.5) +
             labs(title = i))
  }

  #can I facet by particular parameter ranges ?
  #might need to create a categorical var from the range and use that to facet 
  #or subset the data for specific parameter ranges

@@

## Looking more closely at effect of the ratio between the level of resistance to the first insecticide divided by that of the second one. Restricting the x axis to lower values.
```{r, eval=TRUE, echo=FALSE, warning=FALSE}

  i <- 'resist_start_1_div_2'
  color <- "start_freq_allele2"

  print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
           geom_point(shape=3) + 
           facet_wrap( ~ strategy) + 
           #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
           geom_smooth(colour='red', linetype='dashed',size=0.5) +
           labs(title = i) +
           xlim(0,100)
         )

@@

#trying a PRCC analysis using the package sensitivity
```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5}

@@

# pcc(X, y, rank = FALSE, nboot = 0, conf = 0.95)
# 
# ## S3 method for class 'pcc'
# plot(x, ylim = c(-1,1), ...)
# 
# Arguments
# X a data frame containing model input variables
# y a vector containing the responses 
# rank logical. If TRUE, the analysis is done on the ranks.
# nboot the number of bootstrap replicates.
# conf the confidence level of the bootstrap confidence intervals.

@@

```{r, eval=FALSE, echo=FALSE, warning=FALSE, fig.height=3.5}

#trying to find apparent anomaly in PRCC analysis
#what causes correctMixDeployProp to get a high PRCC score for I1 & 2 when its not used in those runs ??

tstI1 <- round(listOutI1$input,3)
tstMix <- round(listOutMix$input,3)
tstDif <- round(tstMix - tstI1,3)

@@

>>>>

---
  title: "paper1_results_figs"
  output: pdf_document
  author: "Andy South"
  date: "`r Sys.Date()`"
---

# Figures for paper 1 
# version 2 - in progress

## Based on 10000 runs with all inputs varying at once which explains the variability at particular input values.

Comparing these insecticide use strategies:
1. Single use
1. Mixture 1 : threshold reached for either insecticide
1. Mixture 2 : threshold reached for either insecticide, switch to sole use of other until it too reaches threshold
1. Mixture 3 : threshold reached for both insecticides in mixture
1. Sequential : sole use of one insecticide, switch to other when threshold reached until it too reaches threshold

## Notable results

1. Time-to-resistance is lowest for single use and for the first insecticide in a mixture. Time-to-resistance is highest for the 2nd insecticide in a mixture.
1. In between these extremes, there is less difference between the alternative strategies of a) sequential and b) mixture switching to sole use. The latter takes slightly longer to reach the threshold.
1. Exposure has the greatest effect on time-to-resistance, but as it has a similar effect across all strategies it has little influence on which strategy is preferable.
1. Effectiveness of the insecticides is the one input that has a notably different effect on the strategies of a) sequential and b) mixture switching to sole use. Below an effectiveness of ~0.55 to 0.6 sequential takes longer, and above this 'mixture switching to sole use' takes longer.

@@

  ## trying with the extended experiment _ex100
  experiment <- 'extended'

  #remember these don't have new exposure column
  load(file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
  load(file=paste0(outFolder,'listOutI1_ex2_10000.rda'))
  load(file=paste0(outFolder,'listOutI2_ex2_10000.rda'))

# ex10000 runs showed effect of selection    
#   load(file=paste0(outFolder,'listOutMix_ex10000.rda'))
#   load(file=paste0(outFolder,'listOutI1_ex10000.rda'))
#   load(file=paste0(outFolder,'listOutI2_ex10000.rda'))

@@

```{r,echo=FALSE,results='hide'}
### calculate times to reach critical points and add them on to the input file 
### for different insecticide strategies, sequential, mix1 and mix2

  #1) sequential : time to resistance for each insecticide in isolation
  #inputs : inAndOutI1, inAndOutI2
  #find time to criticalPoint for insecticide1
  #find time to criticalPoint for insecticide2
  #add together
  resistPointsI1 <- findResistancePoints(listOutI1, locus=1)
  resistPointsI2 <- findResistancePoints(listOutI2, locus=2)  
  resistPointsSeq <- resistPointsI1 + resistPointsI2

@@

@@

```{r, eval=TRUE, echo=FALSE, warning=FALSE} 
# getting data into format for ggplot

@@

  #function to transpose and add strategy column with the passed value
  #also add on the input columns  
  addStrategyColumn <- function(x, value, inputs){
    x <- as.data.frame( t(x) )
    x$strategy <- value

    #inputs <- as.numeric(inputs)
    #transpose inputs
    inputs <- as.data.frame( t(inputs), stringsAsFactors=FALSE )
    #cbind onto the outputs
    x <- cbind(inputs,x)
    x
  }

@@

  #format data in a different way to enable PRCC on difference between sequential & mix2
  #resistPointsMix2_T
  #resistPointsSeq_T

  #columns to remove from first
  indices1 <- which(names(resistPointsMix2_T) %in% c("gen_cP0.1","gen_cP0.5","strategy"))
  #columns to add to 2nd  
  indices2 <- which(names(resistPointsMix2_T) %in% c("gen_cP0.25"))  
  #rename column in 2nd
  tmp <- resistPointsSeq_T[indices2]
  names(tmp) <- "gen_cP0.25seq"

  df_diff <- cbind(resistPointsMix2_T[-indices1], tmp)

  #remove runs that didn't reach a resistance threshold (999), even if just for one strategy
  #>1000 excludes sequential strategy that had a 999 in
  #BEWARE would need to increase 1000 if I increase max generations in the runs
  didntReachThresh <- which( df_diff$gen_cP0.25 > 500 | df_diff$gen_cP0.25seq > 500 )

  #subset by runs not making threshold
  df_diff <- df_diff[-didntReachThresh,]

  #df_diff$mix2_minus_seq0.25 <- df_diff["gen_cP0.25"] - df_diff["gen_cP0.25seq"]
  df_diff["mix2_minus_seq0.25"] <- df_diff["gen_cP0.25"] - df_diff["gen_cP0.25seq"]

  #TODO : later do PRCC on df_diff["mix2_minus_seq0.25"]

@@

# Time-to-resistance across all input values

```{r, eval=TRUE, echo=FALSE, warning=FALSE}

#names_results <- c("time_to_resistance0.5","time_to_resistance0.25","time_to_resistance0.1")
names_results <- c("time_to_resistance0.25")

ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 2"), ]

for(i in names_results)
{
  #x11()

  print( ggplot(ggSubset, aes_string(x='strategy',y=i, color='strategy')) + 
      #ylim(0,450) +
      coord_cartesian( ylim=c(0, 350)) +
      geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = .5 ) +
      scale_x_discrete("Resistance to which insecticide", limits=c("insecticide 1","Mixture 1","Sequential","Mixture 2","Mixture 3"), labels = c("insecticide 1"="Sole use", "Mixture 1"="1st in mix","Sequential"="2nd in sequence", "Mixture 2"="2nd in mix\nswitch to sole", "Mixture 3"="2nd in mix" ) )

  ) #end print
}

# Difference between time-to-resistance for sequential and mixture2

```{r, eval=TRUE, echo=FALSE, warning=FALSE}

#names_results <- c("time_to_resistance0.5","time_to_resistance0.25","time_to_resistance0.1")
# names_results <- c("mix2_minus_seq0.25")
#   
# ggSubset <- df_diff #[ !ggInsOuts$strategy %in% c("insecticide 2"), ]  
#   
# for(i in names_results)
# {
#   #x11()
#   
#   print( ggplot(ggSubset, aes_string(y=i)) + 
#       #ylim(0,450) +
#       #coord_cartesian( ylim=c(0, 350)) +
#       geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = .5 )
# 
#   ) #end print
# }

## How time-to-resistance is effected by input values for each insecticide use strategy

```{r, eval=TRUE, echo=FALSE, fig.height=3.5}

@@

  ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 1","insecticide 2"), ]

@@

    y <- 'time_to_resistance0.25'

    #print( ggplot(ggInsOuts, aes_string(x=i, y=y, colour="strategy")) + 
    print( ggplot(ggSubset, aes_string(x=i, y=y, colour="strategy")) + 
            #points not wanted if 10000
             #geom_point(shape=3, show.legend=FALSE) + 
             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
             geom_smooth(linetype='dashed',size=1) +
             #facet_wrap( ~ strategy) +
             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
             labs(title = i))
  }

@@

## How difference in time-to-resistance for sequential and mixture3 is effected by input values

```{r, eval=TRUE, echo=FALSE, fig.height=3.5}

  ggSubset <- df_diff

  names_inputs <- names(ggSubset)[1:num_inputs]

    y <- "mix2_minus_seq0.25"

    print( ggplot(ggSubset, aes_string(x=i, y=y, colour="gen_cP0.25seq")) + #, colour="strategy")) + 
            #points not wanted if 10000
             geom_point(shape=3, show.legend=FALSE) + 
             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
             geom_smooth(linetype='dashed',size=1) +
             #facet_wrap( ~ strategy) +
             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
             labs(title = i))
  }

@@

# PRCC analysis for time-to-resistance subsetted by each insecticide use strategy
```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5}

@@

# pcc(X, y, rank = FALSE, nboot = 0, conf = 0.95)
# 
# Arguments
# X a data frame containing model input variables
# y a vector containing the responses 
# rank logical. If TRUE, the analysis is done on the ranks.
# nboot the number of bootstrap replicates.
# conf the confidence level of the bootstrap confidence intervals.

@@

## response of resistance thresholds to individual inputs faceted by strategy
```{r, eval=TRUE, echo=FALSE, fig.height=3.5}

  #todo find out to take sample this doesn't work
  #ggSample <- sample(ggInsOuts,500)

    y <- 'time_to_resistance0.25'    
    #y <- 'time_to_resistance0.5'
    color <- 'strategy'
    #coloring by exposure shows its effect even when other params are varying
    #color <- 'exposure'

    print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
             #points not wanted if 10000
             #geom_point(shape=3, show.legend=FALSE) + 
             geom_smooth(colour='red', linetype='dashed',size=0.5, show.legend=FALSE) +
             facet_wrap( ~ strategy) + #, show_guide = FALSE) + 
             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
             labs(title = i))
  }

@@

## Andys questions 
1. In mixtures what causes resistance to the 2nd insecticide to arise more slowly when the first insecticide is removed from the mixture ?
